---
title: "IDS investigation worksheet"
author: "by Team-Name: Terry, Rithvic, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)
library(janitor)
# Add any other libraries here



```


```{r load-data}
# load your data 
sample_submission <- read_csv("data/sample_submission.csv")
test <- read_csv("data/test.csv")
train <- read_csv("data/train.csv")
```

```{r cleaning}
# Cleaning sample_submission
glimpse(sample_submission)
sum(is.na(sample_submission))
#sample_submission is clean by default
```

```{r cleaning}
# Cleaning train
clean_names(train)
train <- train %>%
  mutate(MSSubClass_fctr = as.factor(MSSubClass),
         OverallQual_fctr = as.factor(OverallQual),
         OverallCond_fctr = as.factor(OverallCond),
         ExterQual_fctr = as.factor(ExterQual),
         ExterCond_fctr = as.factor(ExterCond),
         BsmtQual_fctr = as.factor(BsmtQual),
         BsmtCond_fctr = as.factor(BsmtCond),
         BsmtExposure_fctr = as.factor(BsmtExposure),
         BsmtFinType1_fctr = as.factor(BsmtFinType1),
         BsmtFinType2_fctr = as.factor(BsmtFinType2),
         HeatingQC_fctr = as.factor(HeatingQC),
         KitchenQual_fctr = as.factor(KitchenQual),
         Functional_fctr = as.factor(Functional),
         FireplaceQu_fctr = as.factor(FireplaceQu),
         GarageQual_fctr = as.factor(GarageQual),
         GarageFinish_fctr = as.factor(GarageFinish),
         GarageCond_fctr = as.factor(GarageCond),
         PoolQC_fctr = as.factor(PoolQC),
         Fence_fctr = as.factor(Fence))
train %>%
  summarise(across(everything(), ~n_distinct(.x))) %>%
  select(where(~.x == 0))
  
# demonstrates that there are no columns with no arguments, hence this data frame must have no useless columns.
```

```{r cleaning}
# Cleaning test
clean_names(test)
test <- test %>%
 mutate(MSSubClass_fctr = as.factor(MSSubClass),
         OverallQual_fctr = as.factor(OverallQual),
         OverallCond_fctr = as.factor(OverallCond),
         ExterQual_fctr = as.factor(ExterQual),
         ExterCond_fctr = as.factor(ExterCond),
         BsmtQual_fctr = as.factor(BsmtQual),
         BsmtCond_fctr = as.factor(BsmtCond),
         BsmtExposure_fctr = as.factor(BsmtExposure),
         BsmtFinType1_fctr = as.factor(BsmtFinType1),
         BsmtFinType2_fctr = as.factor(BsmtFinType2),
         HeatingQC_fctr = as.factor(HeatingQC),
         KitchenQual_fctr = as.factor(KitchenQual),
         Functional_fctr = as.factor(Functional),
         FireplaceQu_fctr = as.factor(FireplaceQu),
         GarageQual_fctr = as.factor(GarageQual),
         GarageFinish_fctr = as.factor(GarageFinish),
         GarageCond_fctr = as.factor(GarageCond),
         PoolQC_fctr = as.factor(PoolQC),
         Fence_fctr = as.factor(Fence))
test %>%
  summarise(across(everything(), ~n_distinct(.x))) %>%
  select(where(~.x == 0))

# Shows that there are no columns with no arguments, hence this data frame must have no unnecessary columns
```


```{r}
test_price <- test %>%
  full_join(sample_submission, by = 'Id')

test_price %>%
  summarise(avg_price = mean(SalePrice, na.rm = TRUE),
            med_price = median(SalePrice, na.rm = TRUE),
            q25_price = quantile(SalePrice, probs = 0.25, na.rm = TRUE),
            q75_price = quantile(SalePrice, probs = 0.75, na.rm = TRUE),
            sd_price = sd(SalePrice, na.rm = TRUE),
            min_price = min(SalePrice, na.rm = TRUE),
            max_price = max(SalePrice, na.rm = TRUE))

test_price %>%
  ggplot(aes(
    x = `SalePrice`)) +
  geom_histogram(aes(y = after_stat(density)), bins = 20) +
  stat_function(fun = dnorm,
                args = list(mean = 179183.9, sd = 16518.3),
                color = "red",
                linewidth = 1)
```

# We have worked out that the mean price and median price is the same to 3 significance figure accuracy, this suggests that the distribution of sales price for housing could be approximated to a normal distribution and almost symmetrical, implying that there are no outliers in our data.


```{r}

test_price <- test_price %>%
  mutate(
    # log-transform target
    log_saleprice = log(SalePrice),
  )

# Original SalePrice
ggplot(train, aes(x = SalePrice)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of SalePrice")

# Log-transformed SalePrice
ggplot(test_price, aes(x = log_saleprice)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of log(SalePrice)")


```

#The raw SalePrice distribution was highly right-skewed. We applied a natural log transformation, which produced a more symmetric distribution and better supported linear model assumptions.

```{r}
# investigating correlation between SalePrice and LotArea
test_price %>%
  ggplot(aes(
    x = LotArea,
    y = log_saleprice)) +
  geom_point(color = "red") +
  geom_smooth(method = lm) +
  labs(
    title = "Relationship between Property Area and Log Sale Price",
    x = "Total Property Area (sq ft)",
    y = "log(SalePrice)"
  )
```

# From our scatter point plot when adjusted to standard error, we can see that there is a correlation visually between SalePrices and LotArea, predicting a possible dependency. The standard error increases as we approach greater values in LotArea and SalePrices, which suggests possible outliers, but also due to the lack of data.

```{r}
ggplot(test_price, aes(x = GrLivArea, y = log_saleprice)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Relationship between Living Area and Log Sale Price",
    x = "Above-ground living area (sq ft)",
    y = "log(SalePrice)"
  )

```
```{r}
library(tidyverse)

# Compute median per quality level
medians <- test_price %>%
  group_by(OverallQual_fctr) %>%
  summarise(median_price = median(log_saleprice, na.rm = TRUE))

# Jitter plot + median line
ggplot(test_price, aes(x = OverallQual_fctr, y = log_saleprice)) +
  geom_jitter(alpha = 0.25, width = 0.2, color = "steelblue") +
  geom_line(data = medians,
            aes(x = OverallQual_fctr, y = median_price, group = 1),
            color = "red", linewidth = 1.2) +
  geom_point(data = medians,
             aes(x = OverallQual_fctr, y = median_price),
             color = "red", size = 3) +
  labs(
    title = "Log Sale Price by Overall Quality with Median Trend Line",
    x = "Overall Quality (1–10)",
    y = "log(SalePrice)"
  )

test_price %>%
  count(OverallQual_fctr)
```
# In our Log Sale Price Against Median Quality plot, there seems to be an overall trend from quality equals 2 to 10, with an anomoly at 1. By counting how many data points were given for each score, it is clear that the anomoly is a cause due to lack of data, hence unreliable. It is difficult to assume a general correlation for scores with few data points, where scores with much data points has suggested that the SalePrice does not increase, at least not significantly due to the quality assessed.

```{r}
test_price %>%
  mutate(
    Neighborhood = fct_reorder(Neighborhood, log_saleprice, .fun = median)
  ) %>%
  ggplot(aes(x = Neighborhood, y = log_saleprice)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    title = "Log Sale Price Across Neighborhoods",
    x = "Neighborhood (ordered by median price)",
    y = "log(SalePrice)"
  )
test_price %>%
  count(Neighborhood)
```

