---
title: "IDS investigation worksheet"
author: "by Team-Name: Terry, Rithvic, User3, User4 & User5"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
library(readr)
library(janitor)
# Add any other libraries here



```


```{r load-data}
# load your data 
sample_submission <- read_csv("data/sample_submission.csv")
test <- read_csv("data/test.csv")
train <- read_csv("data/train.csv")
```

```{r cleaning}
# Cleaning sample_submission
glimpse(sample_submission)
sum(is.na(sample_submission))
#sample_submission is clean by default
```

```{r cleaning}
# Cleaning train
clean_names(train)
train <- train %>%
  mutate(MSSubClass_fctr = as.factor(MSSubClass),
         OverallQual_fctr = as.factor(OverallQual),
         OverallCond_fctr = as.factor(OverallCond),
         ExterQual_fctr = as.factor(ExterQual),
         ExterCond_fctr = as.factor(ExterCond),
         BsmtQual_fctr = as.factor(BsmtQual),
         BsmtCond_fctr = as.factor(BsmtCond),
         BsmtExposure_fctr = as.factor(BsmtExposure),
         BsmtFinType1_fctr = as.factor(BsmtFinType1),
         BsmtFinType2_fctr = as.factor(BsmtFinType2),
         HeatingQC_fctr = as.factor(HeatingQC),
         KitchenQual_fctr = as.factor(KitchenQual),
         Functional_fctr = as.factor(Functional),
         FireplaceQu_fctr = as.factor(FireplaceQu),
         GarageQual_fctr = as.factor(GarageQual),
         GarageFinish_fctr = as.factor(GarageFinish),
         GarageCond_fctr = as.factor(GarageCond),
         PoolQC_fctr = as.factor(PoolQC),
         Fence_fctr = as.factor(Fence))
train %>%
  summarise(across(everything(), ~n_distinct(.x))) %>%
  select(where(~.x == 0))
  
# demonstrates that there are no columns with no arguments, hence this data frame must have no useless columns.
```

```{r cleaning}
# Cleaning test
clean_names(test)
test <- test %>%
  mutate(MSSubClass_fctr = as.factor(MSSubClass),
         OverallQual_fctr = as.factor(OverallQual),
         OverallCond_fctr = as.factor(OverallCond),
         ExterQual_fctr = as.factor(ExterQual),
         ExterCond_fctr = as.factor(ExterCond),
         BsmtQual_fctr = as.factor(BsmtQual),
         BsmtCond_fctr = as.factor(BsmtCond),
         BsmtExposure_fctr = as.factor(BsmtExposure),
         BsmtFinType1_fctr = as.factor(BsmtFinType1),
         BsmtFinType2_fctr = as.factor(BsmtFinType2),
         HeatingQC_fctr = as.factor(HeatingQC),
         KitchenQual_fctr = as.factor(KitchenQual),
         Functional_fctr = as.factor(Functional),
         FireplaceQu_fctr = as.factor(FireplaceQu),
         GarageQual_fctr = as.factor(GarageQual),
         GarageFinish_fctr = as.factor(GarageFinish),
         GarageCond_fctr = as.factor(GarageCond),
         PoolQC_fctr = as.factor(PoolQC),
         Fence_fctr = as.factor(Fence))
test %>%
  summarise(across(everything(), ~n_distinct(.x))) %>%
  select(where(~.x == 0))

# Shows that there are no columns with no arguments, hence this data frame must have no useless columns
```


```{r}
sample_submission %>%
  ggplot(aes(
    x = SalePrice)) +
  geom_density()
```

```{r}

```

```{r}
# summarizing order type

```
