---
title: "Predicting House Prices in Ames,Iowa"
subtitle: "House features vs Location"
author: "Final Project Duo <br> Terry and Rithvic"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
# Load your data here
sample_submission <- read_csv("data/sample_submission.csv")
test <- read_csv("data/test.csv")
train <- read_csv("data/train.csv")

train <- train %>%
  select(Id,OverallQual,GrLivArea,YearBuilt,Neighborhood,MSZoning,LotFrontage,SalePrice)

train <- train %>%
  mutate(OverallQual = case_when(
    OverallQual > 7 ~ "High Qual",
    OverallQual < 4 ~ "Low Qual",
    OverallQual > 3 & OverallQual < 8  ~ "Med Qual"))

train <- train %>%
  mutate(
  OverallQual = fct_relevel(as.factor(OverallQual), "Low Qual", "Med Qual", "High Qual"))

test <- test %>%
 select(Id,OverallQual,GrLivArea,YearBuilt,Neighborhood,MSZoning,LotFrontage)

test <- test %>%
  mutate(OverallQual = case_when(
    OverallQual > 7 ~ "High Qual",
    OverallQual < 4 ~ "Low Qual",
    OverallQual > 3 & OverallQual < 8  ~ "Med Qual"))

train <- train %>%
  mutate(
    log_saleprice = log(SalePrice),)

train <- train %>%
 mutate(log_avg_price = mean(log_saleprice))

train <- train %>%
  mutate(Mag_Price = case_when(
    log_saleprice > log_avg_price ~ "High",
    log_saleprice < log_avg_price ~ "Low"),
    Mag_Price = fct_relevel(Mag_Price, "Low", "High"))

sample_submission <- sample_submission %>%
  mutate(
    log_saleprice = log(SalePrice),)

sample_submission <- sample_submission %>%
 mutate(log_avg_price = mean(log_saleprice))

sample_submission <- sample_submission %>%
  mutate(Mag_Price = case_when(
    log_saleprice > log_avg_price ~ "High",
    log_saleprice < log_avg_price ~ "Low"),
    Mag_Price = fct_relevel(Mag_Price, "Low", "High"))
```

```{r include=FALSE}

#Background image
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```

### Data Overview:

Our project uses three related datasets derived from the Ames Housing data. After cleaning the data we selected our variables and split them into two groups, Group A (House Features) and Group B (Location Variables).

To support our comparison of house features vs. location factors, we retained the following predictors:

House-related variables: OverallQual, GrLivArea, YearBuilt

Location-related variables: Neighborhood, MSZoning, LotFrontage
```{r, echo=FALSE}

library(knitr)

data.frame(
  Variable = c(
    # Group A — House Features
    "OverallQual", "GrLivArea", "YearBuilt",
    
    # Group B — Location Features
    "Neighborhood", "MSZoning", "LotFrontage"),
  
  Type = c(
    # Group A
    "Categorical",
    "Numeric",
    "Numeric",
    
    # Group B
    "Categorical",
    "Categorical",
    "Numeric"
  ),
  
  Description = c(
    # Group A descriptions
    "Reclassified overall finish and quality of the home",
    "Living area in square feet",
    "Year the house was originally built",
    
    # Group B descriptions
    "Residential neighborhood where the home is located",
    "Zoning classification for land use",
    "Linear feet of street-connected property frontage"
  )
) %>%
  kable()

```
---
We also engineered two new variables:

log_saleprice — the natural log of SalePrice to reduce skew and support linear modelling

```{r, echo=FALSE, fig.show='hold', out.width='48%'}

train <- train %>%
  mutate(
    # log-transform target
    log_saleprice = log(SalePrice),)

# Original SalePrice
ggplot(train, aes(x = SalePrice)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of SalePrice")

# Log-transformed SalePrice
ggplot(train, aes(x = log_saleprice)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of log(SalePrice)")


```

Mag_Price — a binary classification (“High” / “Low”) indicating whether a home’s log-price is above or below the training-set mean

We also recoded OverallQual into three descriptive quality categories (“Low”, “Med”, “High”) for clearer interpretation.



---

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.


```

### Group A (House Features)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width='32%'}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(train, aes(x = factor(OverallQual), y = log_saleprice)) +
  geom_boxplot(fill = "#6AAEDF") +
  theme_minimal() +
  labs(
    title = "log(SalePrice) by Overall Quality",
    x = "Overall Quality",
    y = "log(SalePrice)"
  )

ggplot(train, aes(x = GrLivArea, y = log_saleprice)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  theme_minimal() +
  labs(
    title = "log(SalePrice) vs Living Area",
    x = "GrLivArea (sq ft)",
    y = "log(SalePrice)"
  )

ggplot(train, aes(x = YearBuilt, y = log_saleprice)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "loess", se = FALSE, color = "darkgreen") +
  theme_minimal() +
  labs(
    title = "log(SalePrice) vs Year Built",
    x = "Year Built",
    y = "log(SalePrice)"
  )
```

### Group B (Location Variables)

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold', out.width='32%'}
ggplot(train, aes(
      x = reorder(Neighborhood, log_saleprice, FUN = median),
      y = log_saleprice)) +
  geom_boxplot(outlier.alpha = 0.2) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "log(SalePrice) by Neighborhood",
    x = "Neighborhood (ordered by price)",
    y = "log(SalePrice)"
  )

ggplot(train, aes(x = MSZoning, y = log_saleprice)) +
  geom_boxplot(fill = "#F9A65A") +
  theme_minimal() +
  labs(
    title = "log(SalePrice) by Zoning Classification",
    x = "MSZoning",
    y = "log(SalePrice)"
  )

ggplot(train, aes(x = LotFrontage, y = log_saleprice)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  theme_minimal() +
  labs(
    title = "log(SalePrice) vs Lot Frontage",
    x = "LotFrontage (linear feet)",
    y = "log(SalePrice)"
  )

```

---
#Predictive Modelling 

Since our target variable is whether the price is High or Low, we will compare both the groups above to determine which influences the prices more. As the outcome is binary, we use logistic regression.
```{r}
log_model <- logistic_reg() %>% set_engine("glm")
```
We created two model workflows:
.pull-left[
Group A
```{r}
rec_A <- recipe(
  Mag_Price ~ 
    GrLivArea + OverallQual + YearBuilt,
  data = train
) %>%
  step_dummy(all_nominal_predictors())

wf_A <- workflow() %>%
  add_model(log_model) %>%
  add_recipe(rec_A)
```
]
.pull-right[
Group B
```{r}
rec_B <- recipe(
  Mag_Price ~ 
    Neighborhood + MSZoning + LotFrontage,
  data = train
) %>%
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())

wf_B <- workflow() %>%
  add_model(log_model) %>%
  add_recipe(rec_B)
```
]
---
.pull-left[
### Group A
```{r}
fit_A <- wf_A %>%
  fit(data = train)

pred_A_test <- predict(fit_A, new_data = test, type = "prob") %>%
  bind_cols(test %>% select(Id))

head(pred_A_test)
```
]
.pull-right[
### Group B
```{r, message=FALSE, warning=FALSE}
fit_B <- wf_B %>%
  fit(data = train)

pred_B_test <- predict(fit_B, new_data = test, type = "prob") %>%
  bind_cols(test %>% select(Id))

head(pred_B_test)

```
]
---
## Comparing the Models

```{r, echo=FALSE}
truth_test <- sample_submission %>%
  select(Id, Mag_Price)

roc_data_A <- pred_A_test %>%
  inner_join(truth_test, by = "Id")

roc_data_B <- pred_B_test %>%
  inner_join(truth_test, by = "Id")

# Group A ROC & AUC
roc_A <- roc_data_A %>%
  roc_curve(truth = Mag_Price, .pred_High, event_level = "second")

auc_A <- roc_data_A %>%
  roc_auc(truth = Mag_Price, .pred_High, event_level = "second")

# Group B ROC & AUC
roc_B <- roc_data_B %>%
  roc_curve(truth = Mag_Price, .pred_High, event_level = "second")

auc_B <- roc_data_B %>%
  roc_auc(truth = Mag_Price, .pred_High, event_level = "second")

```


.pull-left[
- AUC of Group A is 0.686
- AUC of Group B is 0.657

While we cannot conclude with confidence that Group A is a superior set of predictors, it is however, in comparison to Group B a marginally better predictor.
]
.pull-right.wide[
```{r warning=FALSE, echo=FALSE, out.width="300%"}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
bind_rows(
  roc_A %>% mutate(group = "Group A (house features)"),
  roc_B %>% mutate(group = "Group B (location)")
) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, colour = group)) +
  geom_path() +
  geom_abline(linetype = "dashed") +
  coord_equal() +
  theme_minimal() +
  labs(
    title = "ROC curves: Group A vs Group B",
    x = "1 - Specificity (False Positive Rate)",
    y = "Sensitivity (True Positive Rate)",
    colour = "Model"
  )
```
]
---
### Why do house features beat location when predicting prices?

- Group A variables are mostly numeric or low-cardinality categorical

- Group A achieves better performance with fewer, stronger predictors.

- Group B required many dummy variables

### Conclusion
Our analysis suggests that both groups can be used to predict house prices to a limited extent.

### References
"House Prices - Advanced Regression Techniques." Link - https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques/overview
